HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2022-02-07 hal_arxiv/hermes_paper.tex
Add the HAL/arxiv version

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[lettersize,journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\usepackage[pdftex]{graphicx}
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath,amssymb}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
 \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
 \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and
% Axel Sommerfeldt. This package may be useful when used in conjunction with
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

\usepackage{booktabs}
\usepackage{url,xcolor}

\input{math_commands.tex}
\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Etienne David\thanks{Etienne David: Heuritech, France.}, Jean Bellot\thanks{Jean Bellot: Heuritech, France.} and Sylvain Le Corff\thanks{Sylvain Le Corff:  Samovar, T\'el\'ecom SudParis, d\'epartement CITI, Institut Polytechnique de Paris, France.}}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell was with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks -
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
%
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}

% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} the first fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \cite{makridakis2018m4}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Hybrid models, Recurrent neural networks, Time series.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \cite{box2015time} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

\begin{figure*}
\centering
%\begin{subfigure}{1.\textwidth}
 % \centering
  \includegraphics[width=\linewidth]{figure/us_female_shoes}
  %\caption{Time series representing a shoes fashion trend for female in The United States.}
 % \label{fig:introexamples:sub1}
%\end{subfigure}
%\begin{subfigure}{1.\textwidth}
  %\centering
  \includegraphics[width=\linewidth]{figure/br_female_texture_verticalstripe}
 % \caption{Time series representing the vertical stipes texture fashion trend for female in Brazil.}
  %\label{fig:introexamples:sub2}
%\end{subfigure}
\caption{Hermes forecast examples. In green the prediction of the TBATS per-time-series predictors. In red the final forecast of our HERMES hybrid model. (Top) Time series representing a shoes fashion trend for females in The United States. (Bottom) Time series representing the vertical stipes texture fashion trend for females in Brazil.}
\label{fig:introexamples}
\end{figure*}

@@

Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and provides very good performance on the fashion dataset, see Table~\ref{tab:metricresults}. Then, to correct the second form of weakness, in addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals associated with the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
\begin{align*}
\fullconcatinput^n_T &= \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}\\
&= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
\end{align*}
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{align}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T}\\
& = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
An illustration of the proposed  model is displayed in Figure~\ref{fig:architecture}.

\begin{figure}
  \centering
    \includegraphics[width=1\linewidth]{figure/HERMES_archi.png}
  \caption{Architecture of the hybrid model with weak signals.}
\label{fig:architecture}
\end{figure}

@@

\begin{figure*}
\centering
%\begin{subfigure}{1.\textwidth}
 % \centering
  \includegraphics[width=1.\linewidth]{figure/cn_female_top_raw}
%  \caption{Time series representing the raw signal of a top fashion trend for female in China}
 % \label{fig:normalization:sub1}
%\end{subfigure}
%\begin{subfigure}{1.\textwidth}
 % \centering
  \includegraphics[width=1.\linewidth]{figure/cn_female_top_norm}
  %\caption{Time series representing the normalized signal of a top fashion trend for female in China.}
  %\label{fig:normalization:sub2}
%\end{subfigure}
\caption{Example of difference between the raw sequence and the normalized one. In this example, we normalize by the deseasonalized global top fashion trend for females in China. (Top) Time series representing the raw signal of a top fashion trend for females in China. (Bottom) Time series representing the normalized signal of a top fashion trend for females in China.}
\label{fig:normalization}
\end{figure*}

\begin{table*}
  \caption{Fashion time series overview. For each couple geozone/category, the table gives the number of trends (Female/Male).}
\label{tab:fashiondataset}
  \centering
  \resizebox{0.85\width}{!}{
  \begin{tabular}{l||lllllllll}
    %hline
    \\
    &  \textbf{Top}  & \textbf{Pants} & \textbf{Short} & \textbf{Skirt} & \textbf{Dress} & \textbf{Coat} & \textbf{Shoes} & \textbf{Color} & \textbf{Texture}  \\
    \hline
    \hline
    \\
\textbf{United States} & 411/208 & 149/112 & 47/22 & 29/- & 20/- & 208/151 & 293/86 & 38/44 & 85/81\\
     \textbf{Europe} & 409/228 &  134/114 & 48/21 & 28/- & 20/- & 211/159 & 303/78 & 41/42 & 87/74\\
     \textbf{Japan} &  403/218 & 136/107 & 49/31 & 28/- &  23/- & 185/149 &  311/78 & 46/42 &  92/65\\
     \textbf{China} &  424/202 & 147/114 & 46/29 & 27/- &  27/- & 178/161 &  310/78 & 41/47 &  88/77\\
     \textbf{Brazil} &  431/222 & 134/117 & 49/27 & 30/- &  28/- & 203/152 & 311/76 & 48/41 & 107/84\\
     \\
     %\hline
     \textbf{Total} & 2078/1078 & 700/564 & 239/130 & 142/- & 118/- & 985/772 & 1528/396 & 214/216 & 459/381\\
    %\hline
  \end{tabular}
  }
\end{table*}

@@

\begin{figure*}
  \centering
    \includegraphics[width=1.\linewidth]{figure/train_eval_test_set}
  \caption{Temporal split for our training process. The three first years define our training set. The fourth year is used as our eval set and the final year is reserved for the test set.}
\label{fig:train_eval_test_set}
\end{figure*}

>>>>

\documentclass[nolayout]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{authblk}
\usepackage{amssymb,amsmath,amsthm}
\usepackage[pdftex]{graphicx}
\usepackage{footnote}
\usepackage{url}
\usepackage{times,dsfont}
\usepackage{fancyhdr}
\usepackage{geometry}
\pagestyle{fancy}

\title{HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series}
\date{}

\author[$\dag$,*]{\'Etienne David}
\author[*]{Jean Bellot}
\author[$\dag$]{Sylvain Le Corff}

\affil[*]{{\small  Samovar, T\'el\'ecom SudParis,  D\'epartement CITI, Institut Polytechnique de Paris, France.}}
\affil[*]{{\small Heuritech, Paris, France.}}

\lhead{\'Etienne David, Jean Bellot, Sylvain Le Corff}
\rhead{Nonparametric general translation HMM}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

@@

\begin{document}

\maketitle

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}}\footnote[2]{\url{https://github.com/etidav/f1fashiondataset}} the first fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviors of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \cite{makridakis2018m4}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

{\bf Keywords:} Hybrid models, Recurrent neural networks, Time series.

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations.
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \cite{box2015time} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure.
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

\begin{figure*}
\centering
  \includegraphics[width=\linewidth]{us_female_shoes}
  \includegraphics[width=\linewidth]{br_female_texture_verticalstripe}
\caption{Hermes forecast examples. In green the prediction of the TBATS per-time-series predictors. In red the final forecast of our HERMES hybrid model. (Top) Time series representing a shoes fashion trend for females in The United States. (Bottom) Time series representing the vertical stipes texture fashion trend for females in Brazil.}
\label{fig:introexamples}
\end{figure*}

@@

Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and provides very good performance on the fashion dataset, see Table~\ref{tab:metricresults}. Then, to correct the second form of weakness, in addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals associated with the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
\begin{align*}
\fullconcatinput^n_T &= \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}\\
&= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
\end{align*}
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:
\begin{align}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T}\\
& = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
An illustration of the proposed  model is displayed in Figure~\ref{fig:architecture}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{HERMES_archi.png}
\caption{Architecture of the hybrid model with weak signals.}
\label{fig:architecture}
\end{figure}

@@

\begin{figure*}
\centering
  \includegraphics[width=1.\linewidth]{cn_female_top_raw}
  \includegraphics[width=1.\linewidth]{cn_female_top_norm}
\caption{Example of difference between the raw sequence and the normalized one. In this example, we normalize by the deseasonalized global top fashion trend for females in China. (Top) Time series representing the raw signal of a top fashion trend for females in China. (Bottom) Time series representing the normalized signal of a top fashion trend for females in China.}
\label{fig:normalization}
\end{figure*}

\begin{table*}
  \caption{Fashion time series overview. For each couple geozone/category, the table gives the number of trends (Female/Male).}
\label{tab:fashiondataset}
  \centering
  \resizebox{0.85\width}{!}{
  \begin{tabular}{l||lllllllll}
    \\
    &  \textbf{Top}  & \textbf{Pants} & \textbf{Short} & \textbf{Skirt} & \textbf{Dress} & \textbf{Coat} & \textbf{Shoes} & \textbf{Color} & \textbf{Texture}  \\
    \hline
    \hline
    \\
\textbf{United States} & 411/208 & 149/112 & 47/22 & 29/- & 20/- & 208/151 & 293/86 & 38/44 & 85/81\\
     \textbf{Europe} & 409/228 &  134/114 & 48/21 & 28/- & 20/- & 211/159 & 303/78 & 41/42 & 87/74\\
     \textbf{Japan} &  403/218 & 136/107 & 49/31 & 28/- &  23/- & 185/149 &  311/78 & 46/42 &  92/65\\
     \textbf{China} &  424/202 & 147/114 & 46/29 & 27/- &  27/- & 178/161 &  310/78 & 41/47 &  88/77\\
     \textbf{Brazil} &  431/222 & 134/117 & 49/27 & 30/- &  28/- & 203/152 & 311/76 & 48/41 & 107/84\\
     \\
     \textbf{Total} & 2078/1078 & 700/564 & 239/130 & 142/- & 118/- & 985/772 & 1528/396 & 214/216 & 459/381\\
  \end{tabular}
  }
\end{table*}

@@

\begin{figure*}
  \centering
    \includegraphics[width=1.\linewidth]{train_eval_test_set}
  \caption{Temporal split for our training process. The three first years define our training set. The fourth year is used as our eval set and the final year is reserved for the test set.}
\label{fig:train_eval_test_set}
\end{figure*}

\begin{figure}
  \centering
    \includegraphics[width=0.8\linewidth]{lstm_archi}
  \caption{Architecture of the RNN corrector part of the HERMES framework. The same architecture is used in the \textit{lstm} benchmark model.}
\label{fig:rnn_architecture}
\end{figure}
